Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-21 intro.tex
2012.05.29 (Vassilis) Monolithic file broken down to multiple smaller ones (per section) for better parallel contribution. Additional material on more recent algorithms for unstructured networks. Now also covering the BitTorrent protocol. Merged Huseyin's figures from previous contribution and code refactored for better referencing. Fixed bibliography bug (references appearing as ?).

%\documentclass[a4paper,10pt]{article}
\documentclass[acmcsur,acmnow]{acmtrans2m}
%\documentclass[acmtocl,acmnow]{acmtrans2m}

%\acmVolume{V}
%\acmNumber{N}
%\acmYear{YY}
%\acmMonth{Month}

% packages
\usepackage[utf-8]{inputenc}
\usepackage{paralist}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{pdflscape}

% definitions
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newdef{definition}[theorem]{Definition}
\newdef{remark}[theorem]{Remark}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{definition}[theorem]{Definition}

%\textwidth      6.5in
%\textheight     9.2in
%\oddsidemargin  -0.5cm
%\evensidemargin -0.5cm
%\parindent      0pt              
%\topmargin      -0.3in
%\headsep        20pt
%
%\parskip        5pt
%\parindent      10pt

\markboth{}{}{}

\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}

\author{
VASSILIS S. MOUSTAKAS\\\emph{National and Kapodistrian University of Athens}\\
H\"USEYIN AKCAN\\\emph{Izmir University of Economics}\\
MEMA ROUSSOPOULOS\\\emph{National and Kapodistrian University of Athens}\\
AND\\
ALEX DELIS\\\emph{National and Kapodistrian University of Athens}\\
%  Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
%  $^1$Department of Informatics and Telecommunications,\\
%  National and Kapodistrian University of Athens\\
%  \mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
%  $^2$Department of Software Engineering,\\
%  Izmir University of Economics, Izmir, Turkey \\
%  \mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
}

\begin{abstract}
Peer-to-peer (P2P) systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize a plethora of ever-wanted exotic features like anonymity, high availability and robustness, load balancing, quality-of-service (QoS) at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay and the physical underlying network, a problem that imposes a huge amount of unnecessary stress to the network resources and, thus, impede applications from scaling. In this survey we investigate on the recent research developments towards the alleviation of this problem, in both unstructured and structured overlay architectures. We visit the various algorithms available in the lietrature, focusing on the presentation of how each of them tries to tackle the problem. Additionally, we provide a finer grained categorisation of the approaches based on the generic type of methodology they use in order to accomplish this goal.
% TODO: Review this after the comparison sections are fixed.
Finally we discuss the advantages, disadvantages and novelties introduced by the algotithms and offer pictorial comparisons for them in order to further strengthen and organise understanding of the information and knowledge cited, in an attempt to make this survey a one-stop-shop reference for anyone interested in the topic, from researchers and students, to system architects and developers.
\end{abstract}

\category{}{}{}

\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}

\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer architecture, overlay network }

\begin{document}

\begin{bottomstuff}
SAY SOMETHING COOL HERE...
\end{bottomstuff}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The abstraction of the overlay network has been proposed as a way to implement efficient, fully distributed, application-layer services. For example, backing the routing of messages to destinations that are not known in advance or building one-to-many addressing schemes
% TODO find citation
%[Narada, Adobe's Real Time Media Flow Protocol (RTMFP)]
that would provide higher quality-of-service (QoS) guarantees. \emph{IP multicast}, as well as other proposals such as \emph{IntServ} and \emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance (yet?), largely due to the fact they require support from the physical infrastructure. On the other hand, an overlay network can be deployed on end-systems running the overlay protocol software, without cooperation from the ISPs.
%Academic research includes \cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Overlay networks on top of the best-effort IP layer, formed a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for short. P2P systems consist of a number of networked computers called \emph{peers}, where each, such, \emph{peer} could act both as a resource provider and a resource consumer. This changed the traditional \emph{client-server} model dominating the Internet and lead to the introduction of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends the notions of server and client to denote the twofold role of the P2P network participants. Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as more and more participants join the system without requiring additional expenditure for infrastructure\footnote{
        Unfortunatelly, there has been observed mitigation of this self-scaling property by the undesirable behaviour of participating users, usually answered in popular file-sharing systems; from \emph{free-riding} \cite{saroiu_measurefileshare_2002, adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005} to the distribution of illegal content and/or other \emph{socio-technical} issues \cite{hughes_socp2p_2008}.
}.
P2p network systems gained extra attention whithin the distributed systems super-set, circa 2000, due to the fact that they succesfully supported popular applications that offered file sharing among vast numbers of participating users. But their development didn't stay still. On the contrary, during the last decade, research on P2P systems was constant and fruitfull generating a wide range of popular protocols, networks and applications. As P2P systems evolved three main architectures have been emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

@@

The centralized approach has been eventually replaced by architectures that distributed both search and download capabilities (see Figures~\ref{figure:p2p-archs:unstructured-full} and \ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. For this reason they are also refered to as \emph{unstructured}. The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers or super-peers to propagate queries
%\footnote{The life of each query message is constrained by an integer, refered to as \emph{time-to-live (TTL)}. When a querry is received by a (super-) peer, this value is decreased by one unit. Then, the receiver, back-propagates its results and if the TTL has not reached zero the node forwards the query message to all its neighbours except from the one it received it from.}
.

Recently, \emph{decentralized structured} schemes have been proposed (see Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising infrastructure for large-scale P2P applications \cite{ratnasamy_can_2001, stoica_chord_2001, antony_pastry_2001, zhao_tapestry_2001, maymounkov_kademlia_2002, rgrk_bamboo_2004}. They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism. Even though unstructured networks became highly popular among file sharing applications, they do have a certain limitation in locating rare objects. The DHT approach of structured networks on the other hand, provides a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located, even in the case where only a single copy exists in the system.  This means $O \left ( log n \right )$ hops, compared to unstructured networks that require $O \left ( n \right )$ to reliably locate a specific object\footnote{Unfortunatelly this efficiency does not come without cost. Structured overlays realise what is called object identifier location as opposed to content-based retrieval of unstructured ones. This means that the later can support more versatile object location mechanisms like keyword and partial matching searches.}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Applications and Implications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The most successful incarnation of the centrilised approach is without a doubt the popular file-sharing application \emph{Napster}\footnote{Pioneering, though, was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare processing power of participants to search the universe for the existence of extra terrestrial intelligence!}
% TODO find citation for Napster again
\cite{}.
Napster maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever'' that toped some 26 million users and 80 million songs \cite{jmm_naptopusage_2001}.

Napster was ultimatelly brought down after Recording Industry Association of America (RIAA) filed a lawsuit for copyright infringement of the intellectual property of artists whose music was distributed by the network\footnote{
      The sueing spree was initiated by heavy metal giants \emph{Metallica} that feared Napster was undermining record sales. Opposed opinions instantly emerged  with best example the british alternative rock band \emph{Radiohead} whose album \emph{Kid A} had also been leaked to Napster three months before reaching music store shelves \cite{rm_radioheadkida_2000}. Despite music critics labeling the album ``experimental'' and ``challenging'' (it had no official singles nor music videos for any of its songs), in October 2000 Kid A captured the number one spot on the Billboard 200 sales chart in its debut week. Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster, Radiohead had never hit the top20 in the US before.
}
\cite{wiki_amnaptrial_2001}. This triggered a shift towards decentralised networks like the classic \emph{Gnutella} and, its ``rival'' at the time, \emph{FastTrack}. Gnutella, in its first realisation (v.0.4 \cite{gnutellav0.4}), was a fully decentralised approach. Unfortunatelly, network resourses tended to be overwhelmed by message flooding and if queries were constrained to live for some (small) number of hops, this ultimatelly resulted to a shrinked search scope, leading to a death spiral that prevented the network from scaling. The solution came with a slightly more structured network by introducing the notions of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}). With \emph{Limewire}, the most widely used software that implemented the protocol, the network was highly successful and was reported to have reached connecting more that 3 million users concurrently by the end of 2006 \cite{rsr_gnutellaevol_2006}. Fasttrack and its implementing software KaZaA \cite{kazaa}, on the other hand, went with the hierarchical approach (\emph{super-nodes} in their context) from the very begining, dynamically assigning indexing functions to special peers in order to increase search performace and scaling potential. FastTrack network had an average of 3 million concurrently connected nodes in 2003 \cite{lkr_kazaameasure_2005}. It is estimated that the total number of users was greater than that of Napster at its peak. With Limewire discontinued/shutdown in 2010 and its successor, \emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well signal the end of the once-mighty Gnutella network. KaZaA continues as a propriatery music store solution. In the meanwhile, Fastrack design team had already introduced a VoIP and text messaging application based on the design principles of KaZaA's hierarchical P2P nature, that was meant to change how people communicate around the world. This app was no other than Skype \cite{skype}. Despite its wide usage, little is known for the technical aspects of its overlay network, mainly due to its proprietary nature. Studies, though, confirm that it shares strong structural similarities to its file-sharing predecessor (ordinary hosts and super-nodes in Skype's terminology) but they significantly differentiate in terms of traffic behaviour \cite{g_voipskypesec_2005, gdj_skypestudy_2006}. Skype was introduced in 2003 and from then on it witnessed frenzy-rythmed increase to its popularity, with more that 650 milion registered users in 2011  \cite{skypetotalusers} and reported to have broken the barrier of 35 million skypers concurrently online on the the 5th of March 2012 \cite{skypesymusers}. On October 2011 Microsoft Corp. and Skype Global S.à.r.l. anounnced the completion of Microsoft's acquisition of Skype for \$$8.5$ billion \cite{skypemicrosoft}.

File sharing was the most popular application area for structured architectures as well. Overnet
% TODO find citation
\cite{},
for example, was created for sharing large files (e.g., movies and CD images). It implemented the Kademlia tree-like partitioning algorithm. Other implementations of Kademlia are the Kad Network (eMule/aMule  clients), BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent, BitComet, Transmission and BitSpirit all share compatibility with the Mainline DHT which realises the Kademlia protocol) and the Gnutella DHT (first implemented by LimeWire as the Mojito DHT). Lately, RetroShare \cite{retroshare} is the new trend for file-sharing which incorporates an implementation of the Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of a darknet as well. But the potential of structured architectures was not exhausted only to the above. Structured P2P networks became popular in distributed systems because they could scale smoothly and have fast and reliable object location mechanisms. Thus, research and development was very active and productive resulting in a variety of diverse applications such like distributed search engines \cite{yaci}, distributed data storage systems \cite{kbc_oceanstore_2000, bdet_fsdfs_2000, dkkms_cfs_2001, dr_pastutility_2001, abc_farsite_2002, mmfc_ivy_2002, arla, agebh_dks_2003}, Web caches, archives, and publishing systems \cite{ird_squirrel_2002, bags_youserv_2002, wrc_publius_2000, wm_tangler_2001}, messaging applications \cite{threedegrees}, event notification infrastructures \cite{rkcd_scribe_2001, cdkr_scribe_2002, agebh_dks_2003}, naming services \cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.

%%%%%%%% NEW TRENDS %%%%%%%%
%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in response to the continued push for more anti-piracy tools, the final Pirate Bay verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA. This leads to users trying to find new solutions for anonymous, uncensored, decentralised file sharing solutions like Freenet, Tribler and RetroShare where connections are made only between trusted peers — sometimes called "friends" and the respective formed network F2F.

@@

The huge numbers of users, for the various applications we presented above, reveals the tremendous popularity P2P architectures have enjoyed. This, though, does not come without any toll. Measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that the P2P traffic contributes the largest portion of the overall Internet traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004}. Even with later studies, this observation still holds on. Using deep packet inspection techniques deployed at selected customer sites \emph{Ipoque} and \emph{CacheLogic} were able to track down application usage \cite{cachelogic, ipoque2007, ipoque2009}. CacheLogic reports that more than 60\% of the traffic is identified as P2P, with Ipoque verifying this observation reporting similar results for 2007, while observing some small decline for 2008/2009 (but still remaining dominant compared to other applications). Apart from its participation in overall usage terms, P2P is already a huge contibutor to the overall data exchanged across the Internet\footnote{
    Mainly because file-sharing P2P is used to exchange big files of HD video or high sampled audio content compared to simple Web content distributed via, for example, HTTP.
}.
\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes per month and the potential is to be quintubled by the end of 2012 reaching almost 8 petabytes per month.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Topology Mismatch Problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
One of the major issues that defines the efficiency of an overlay network is the mapping of the overlay structure to the underlying infrastructure. Remember, in Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in different networks. The top layer represents the overlay interconnection formed by these nodes at a higher level. Link connections there, can change as needed by the running environment, without any particular constraint by the underlying physical topology. Let's have a closer look with the help of a simple example. Assume nodes $A$, $B$, $C$ and $D$ are connected through the physical network shown in Figure~\ref{figure:phys}, where the network costs are given in milliseconds, and peer $A$ sends a message to peer $D$.

>>>>

\section{Introduction}
\label{section:intro}

@@

The abstraction of the overlay network has been proposed as a way to implement
efficient, fully distributed, application-layer services. For example, backing
the routing of messages to destinations that are not known in advance or
building one-to-many addressing schemes
% TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
\cite{} that would provide higher quality-of-service (QoS) guarantees. \emph{IP
multicast}, as well as other proposals such as \emph{IntServ} and
\emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance
(yet?), largely due to the fact they require support from the physical
infrastructure. On the other hand, an overlay network can be deployed on
end-systems running the overlay protocol software, without cooperation from the
ISPs.
%Academic research includes
%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Overlay networks on top of the best-effort IP layer, formed a widely ranged
family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for
short. P2P systems consist of a number of networked computers called
\emph{peers}, where each, such, \emph{peer} could act both as a resource
provider and a resource consumer. This changed the traditional
\emph{client-server} model dominating the Internet and lead to the introduction
of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends
the notions of server and client to denote the twofold role of the P2P network
participants. Such serverless systems, proved to be able to achieve outstanding
aggregate resource capacities as more and more participants join the system
without requiring additional expenditure for infrastructure\footnote{
  Unfortunatelly, there has been observed mitigation of this self-scaling
  property by the undesirable behaviour of participating users, usually answered
  in popular file-sharing systems; from \emph{free-riding}

\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005}
to the distribution of illegal content and/or other \emph{socio-technical}
issues \cite{hughes_socp2p_2008}.
}.
P2p network systems gained extra attention whithin the distributed systems
super-set, circa 2000, due to the fact that they succesfully supported popular
applications that offered file sharing among vast numbers of participating
users. But their development didn't stay still. On the contrary, during the last
decade, research on P2P systems was constant and fruitfull generating a wide
range of popular protocols, networks and applications. As P2P systems evolved
three main architectures have been emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

@@

The centralized approach has been eventually replaced by architectures that
distributed both search and download capabilities (see
Figures~\ref{figure:p2p-archs:unstructured-full} and
\ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized}
architectures, file placement is random, which means there is no correlation
with the network topology whatsoever \cite{yang_improvep2psearch_2002}. For this
reason they are also refered to as \emph{unstructured}. The most important
properties of such systems are that they support inherent heterogeneity of
peers, are highly resilient to peer failures, and incur low maintenance overhead
at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In
the literature, they are also known as \emph{broadcast-based} systems, because
they use \emph{message flooding} among peers or super-peers to propagate
queries.
%\footnote{The life of each query message is constrained by an integer, refered
%to as \emph{time-to-live (TTL)}. When a querry is received by a (super-) peer,
%this value is decreased by one unit. Then, the receiver, back-propagates its
%results and if the TTL has not reached zero the node forwards the query message
%to all its neighbours except from the one it received it from.}

Recently, \emph{decentralized structured} schemes have been proposed (see
Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising
infrastructure for large-scale P2P applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
through a deterministic mechanism. Even though unstructured networks became
highly popular among file sharing applications, they do have a certain
limitation in locating rare objects. The DHT approach of structured networks on
the other hand, provides a guaranteed bound on the number of overlay routing
hops that have to be taken in order for an object to be located, even in the
case where only a single copy exists in the system.  This means
$O \left ( log n \right )$ hops, compared to unstructured networks that require
$O \left ( n \right )$ to reliably locate a specific object\footnote{
  Unfortunatelly this efficiency does not come without cost. Structured overlays
  realise what is called object identifier location as opposed to content-based
  retrieval of unstructured ones. This means that the later can support more
  versatile object location mechanisms like keyword and partial matching
  searches.}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Applications and Implications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The most successful incarnation of the centrilised approach is without a doubt
the popular file-sharing application \emph{Napster}\footnote{Pioneering, though,
was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare
processing power of participants to search the universe for the existence of
extra terrestrial intelligence!}
% TODO find citation for Napster again
\cite{}.
Napster maintained a \emph{central index server} based on file lists provided by
participating peers. The central index server was queried by the users and it
returned pointers to the actual content. Thus, by centralizing search while
distributing downloads, Napster achieved a highly functional design that was
widely acknowledged, at the time, as ``the fastest growing Internet application
ever'' that toped some 26 million users and 80 million songs
\cite{jmm_naptopusage_2001}.

Napster was ultimatelly brought down after Recording Industry Association of
America (RIAA) filed a lawsuit for copyright infringement of the intellectual
property of artists whose music was distributed by the network\footnote{
  The sueing spree was initiated by heavy metal giants \emph{Metallica} that
  feared Napster was undermining record sales. Opposed opinions instantly
  emerged with best example the british alternative rock band \emph{Radiohead}
  whose album \emph{Kid A} had also been leaked to Napster three months before
  reaching music store shelves \cite{rm_radioheadkida_2000}. Despite music
  critics labeling the album ``experimental'' and ``challenging'' (it had no
  official singles nor music videos for any of its songs), in October 2000 Kid A
  captured the number one spot on the Billboard 200 sales chart in its debut
  week. Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster,
  Radiohead had never hit the top20 in the US before.
}
\cite{wiki_amnaptrial_2001}. This triggered a shift towards decentralised
networks like the classic \emph{Gnutella} and, its ``rival'' at the time,
\emph{FastTrack}. Gnutella, in its first realisation (v.0.4 \cite{gnutellav04}),
was a fully decentralised approach. Unfortunatelly, network resourses tended to
be overwhelmed by message flooding and if queries were constrained to live for
some (small) number of hops, this ultimatelly resulted to a shrinked search
scope, leading to a death spiral that prevented the network from scaling. The
solution came with a slightly more structured network by introducing the notions
of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}). With
\emph{Limewire}, the most widely used software that implemented the protocol,
the network was highly successful and was reported to have reached connecting
more that 3 million users concurrently by the end of 2006
\cite{rsr_gnutellaevol_2006}. Fasttrack and its implementing software KaZaA
\cite{kazaa}, on the other hand, went with the hierarchical approach
(\emph{super-nodes} in their context) from the very begining, dynamically
assigning indexing functions to special peers in order to increase search
performace and scaling potential. FastTrack network had an average of 3 million
concurrently connected nodes in 2003 \cite{lkxr_kazaameasure_2005}. It is
estimated that the total number of users was greater than that of Napster at its
peak. With Limewire discontinued/shutdown in 2010 and its successor,
\emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well
signal the end of the once-mighty Gnutella network. KaZaA continues as a
propriatery music store solution. In the meanwhile, Fastrack design team had
already introduced a VoIP and text messaging application based on the design
principles of KaZaA's hierarchical P2P nature, that was meant to change how
people communicate around the world. This app was no other than Skype
\cite{skype}. Despite its wide usage, little is known for the technical aspects
of its overlay network, mainly due to its proprietary nature. Studies, though,
confirm that it shares strong structural similarities to its file-sharing
predecessor (ordinary hosts and super-nodes in Skype's terminology) but they
significantly differentiate in terms of traffic behaviour
\cite{g_voipskypesec_2005,gdj_skypestudy_2006}. Skype was introduced in 2003
and from then on it witnessed frenzy-rythmed increase to its popularity, with
more that 650 milion registered users in 2011 \cite{skypetotalusers} and
reported to have broken the barrier of 35 million skypers concurrently online on
the the 5th of March 2012 \cite{skypesymusers}. On October 2011 Microsoft Corp.
and Skype Global S.à.r.l. anounnced the completion of Microsoft's acquisition of
Skype for \$$8.5$ billion \cite{skypemicrosoft}.

File sharing was the most popular application area for structured architectures
as well. Overnet
% TODO find citation for Overnet
\cite{},
for example, was created for sharing large files (e.g., movies and CD images).
It implemented the Kademlia tree-like partitioning algorithm. Other
implementations of Kademlia are the Kad Network (eMule/aMule  clients),
BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent,
BitComet, Transmission and BitSpirit all share compatibility with the Mainline
DHT which realises the Kademlia protocol) and the Gnutella DHT (first
implemented by LimeWire as the Mojito DHT). Lately, RetroShare \cite{retroshare}
is the new trend for file-sharing which incorporates an implementation of the
Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of
a darknet as well. But the potential of structured architectures was not
exhausted only to the above. Structured P2P networks became popular in
distributed systems because they could scale smoothly and have fast and reliable
object location mechanisms. Thus, research and development was very active and
productive resulting in a variety of diverse applications such like distributed
search engines \cite{yaci}, distributed data storage systems
\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging applications \cite{threedegrees}, event notification infrastructures
\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
\cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and
lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.

%%%%%%%% NEW TRENDS %%%%%%%%
%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in
%response to the continued push for more anti-piracy tools, the final Pirate Bay
%verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA. This
%leads to users trying to find new solutions for anonymous, uncensored,
%decentralised file sharing solutions like Freenet, Tribler and RetroShare where
%connections are made only between trusted peers — sometimes called "friends"
%and the respective formed network F2F.

@@

The huge numbers of users, for the various applications we presented above,
reveals the tremendous popularity P2P architectures have enjoyed. This, though,
does not come without any toll. Measurements on some popular systems, such as
FastTrack, Gnutella, DirectConnect and BitTorrent have shown that the P2P
traffic contributes the largest portion of the overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}.
Even with later studies, this observation still holds on. Using deep packet
inspection techniques deployed at selected customer sites \emph{Ipoque} and
\emph{CacheLogic} were able to track down application usage \cite{cachelogic,ipoque2007,ipoque2009}.
CacheLogic reports that more than 60\% of the traffic is identified as P2P, with
Ipoque verifying this observation reporting similar results for 2007, while
observing some small decline for 2008/2009 (but still remaining dominant
compared to other applications). Apart from its participation in overall usage
terms, P2P is already a huge contibutor to the overall data exchanged across the
Internet\footnote{
  Mainly because file-sharing P2P is used to exchange big files of HD video or
  high sampled audio content compared to simple Web content distributed via, for
  example, HTTP.
}.
\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes
per month and the potential is to be quintubled by the end of 2012 reaching
almost 8 petabytes per month.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Topology Mismatch Problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

One of the major issues that defines the efficiency of an overlay network is the
mapping of the overlay structure to the underlying infrastructure. Remember, in
Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network,
while $C$ and $D$ are in different networks. The top layer represents the
overlay interconnection formed by these nodes at a higher level. Link
connections there, can change as needed by the running environment, without any
particular constraint by the underlying physical topology. Let's have a closer
look with the help of a simple example. Assume nodes $A$, $B$, $C$ and $D$ are
connected through the physical network shown in Figure~\ref{figure:phys}, where
the network costs are given in milliseconds, and peer $A$ sends a message to
peer $D$.
